Synthesising Concise Termination and Non-Termination Proofs for Bit-Vector Programs

### 2014-07-08 papers/termination/paper.tex
Kolmogorov

To give a flavour for the nature of the ``low Kolmogorov complexity'' assumption, 
we observe that linear functions have low Kolmogorov complexity (since a linear function can
be computed with at most 1 instruction and 1 multiplication for each program variable).
Conversely, the probability that a random program of length $l$ computes
a linear function is $O(2^{-l})$, since otherwise we'd need to have an exponential
number of programs computing the same function.

Therefore the number of nonlinear functions $f$ with $K(f) \leq \kappa$
grows exponentially with $\kappa$, while the number of linear functions
grows only linearly.  This shows that assuming the existence of a linear
ranking function is much stronger than assuming the existence of a short
termination proof.  In addition to being a much weaker assumption,
we argue that low Kolmogorov complexity is a more natural assumption than
linearity.  Humans tend to write programs they can understand.  It's hard
to quantify exactly what ``understandable'' means, but we think that
``having a short proof'' is closer to the mark than ``having a linear proof''.

>>>>

To give a flavour for the nature of the ``low Kolmogorov complexity'' assumption, 
we observe that linear functions have low Kolmogorov complexity (since a linear function can
be computed with at most 1 instruction and 1 multiplication for each program variable),
so the number of linear functions $f$ with $K(f) \leq \kappa$ grows linearly with $\kappa$.
Conversely, the number of functions $f$ with $K(f) \leq \kappa$
grows exponentially with $\kappa$.  This shows that assuming the existence of a linear
ranking function is much stronger than assuming the existence of a short
termination proof.  In addition to being a much weaker assumption,
we argue that low Kolmogorov complexity is a more natural assumption than
linearity.  Humans tend to write programs they can understand.  It's hard
to quantify exactly what ``understandable'' means, but we think that
``having a short proof'' is closer to the mark than ``having a linear proof''.
